On the Dynamics of Anthropomorphism in Robotics

### 2013-07-08 dynamics-anthropomorphism.tex
Use hyperref package to generate PDF index
[wip] First pass on Anthropomorphism during Stabilization Phase

@@

Anthropomorphism is quantified by a \emph{normalized level of anthropomorphic
effects}: because anthropomorphic effects are not quantified on an absolute
scale, we present them as a normalized value, that spans from a minimum (no
anthropomorphic effects) to a maximum (corresponding to the novelty effect peak
on Figure~\ref{fig:dynamics}). The actual maximum value of anthropomorphic
effects depends on each unique combinations of human, robot and several other
factors we introduce below, and thus varies. The general \emph{shape} of the
model remains however the same and depicts the level of anthropomorphism over
time, \ie the general dynamics of anthropomorphism.

@@

This is however to modulate depending on the real and percieved reasons of the
unexpected behavior.  Section~\ref{sec:disruptive} discusses more in depth the
mental correlates of disruptive behaviors. It proposes a typology that relates
them to the anthropomorphism they induce. It may increase if the user perceives
(rightfully or not) a form of intentionality, as well as decrease if the
unexpected behavior is perceived as a failure.

@@

@@

\fxnote{Note that ICA and SLA are NOT correlated}

>>>>

\usepackage[bookmarks]{hyperref}

@@

Anthropomorphism is quantified by a \emph{normalized level of anthropomorphic
effects}: because anthropomorphic effects are not quantified on an absolute
scale, we present them as a normalized value, that spans from a minimum (no
anthropomorphic effects) to a maximum (corresponding to the novelty effect peak
on Figure~\ref{fig:dynamics}). The actual maximum value of anthropomorphic
effects depends on each unique combinations of human, robot and several other
factors we introduce below, and thus varies. The general \emph{shape} of the
model remains however the same and depicts the evolution of anthropomorphism over
time, \ie the general dynamics of anthropomorphism.

@@

This is however to be modulated depending on the real and percieved reasons of the
unexpected behavior.  Section~\ref{sec:disruptive} discusses more in depth the
mental correlates of disruptive behaviors. It proposes a typology that relates
them to the anthropomorphism they induce. It may increase if the user perceives
(rightfully or not) a form of intentionality, as well as decrease if the
unexpected behavior is perceived as a failure.

@@

If, after the familiarization phase, no anthropomorphic effects are observed
anymore, the \emph{Stabilized Level of Anthropomorphism} is zero. This can either be interpreted as the user
considering the robot as merely a tool, or as a consequence of no interactions
taking place anymore with the robot (robot not used)\fxwarning{reference to the
roomba experiment}.

@@

The \emph{Stabilized Level of Anthropomorphism} (SLA on
Figure~\ref{fig:dynamics}) describes the long-term lasting level of
anthropomorphism.  Like the ICA, the SLA has a unique value per triplet
(\emph{human}, \emph{robot}, \emph{context of interaction}).

We suggested that the ICA is built on three factors (section~\ref{sec:ica}):
user's \emph{personality}, robot's \emph{design} and interaction \emph{purpose}
(or \emph{interaction context}). The user's personality and the context of use
do also influence the SLA. In particular, it appears that the user's level of
acquaintance with technologies plays an important role in long-term tendency to
anthropomophize (people more familiar with technology understand, and hence
predict, better the behavior of the robot, which in turn lead them more
frequently to ultimately consider the robot as a simple tool).

The robot's design, on the other hand, plays a more subtle role: lasting
anthropomorphic effects have been observed on non-anthropomorphic robots (like
the iRobot Roomba or the military iRobot PackBot\footnote{Rodney Brooks has
reported in keynotes that occasionaly soldiers would give a name to
\emph{their} PackBot and require it to be repared instead of being replaced by
another one in case of incident.}), and on the contrary, anthropomorphic
designs can lead to higher expectation deceptions, resulting in the robot not
being used anymore.

Note that the \emph{Initial Capital of Anthropomorphism} and the
\emph{Stabilized Level of Anthropomorphism} are generally not correlated: one
individual may have high expectations (high ICA), and get disappointed by the
actual abilities of the robots, down to considering it as a simple machine/tool
(low SLA), while another one with the same high ICA may, for instance, create
lasting affective bonds with the robot, and keep anthropomorphizing it (high
SLA).
